---
title: "Trait Changes in Stressed Cyanobacteria Monocultures"
author: "Mark Holmes"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message=FALSE)
library('tidyverse')
library('flowClust')
library('gauseR')
library('deSolve')
library('scales')
```

# Introduction
As part of ongoing experiments seeking to investigate the effects of intraspecific trait variability (ITV) on community resistance to environmental change drivers (DIVERCE) project, we have performed experiments on monoculture populations of the study species. 

# Methods
## Experiment
In this subset of experiments, we cultured 4 strains of *Synechococcus* cyanobacteria in control conditions and with an added herbicide pollutant (atrazine). 

The control conditions consisted of medium, light, etc...

The cultures were inoculated at...

The strains were left to grow for 7 days in either control or stressed conditions...

Atrazine concentration

## Analysis
We used the `GauseR` package to determine the Lotka-Volterra growth parameters of the strains

# Results
## Species growth
Below, we visualise growth of the the strain populations over time. Each point is the average value of three replicates.

```{r, import-data, fig.width=10, fig.height=8, fig.cap="Strain population change over time during 7 days of growth"}
# Load data
mono.data <- read.csv('cyanobacteria/mono_data.csv') %>%
  group_by(strain, treat) %>% # group by strain ID and treatment
  mutate(date.time = date.time %>% as.POSIXct %>% as.numeric, # convert to good format
         date.time = (date.time - min(date.time)) / 3600,  # in hours 
         #
         # ISSUE: THIS ONLY WORKS WHEN NOT IN HOURS??
         #
         strain = as.factor(strain)) %>%
   arrange(date.time) %>% # order by sampling time
  mutate()
```


## Filtering lag phases out using Kremer's [growthTools](https://github.com/ctkremer/growthTools/)

```{r, filter-lag}
library('growthTools')

lag.sat.pars <- mono.data %>% 
  group_by(strain, treat) %>%
  summarise(mods = list(get.growth.rate(
    date.time, log(population.mean), id = paste(strain, treat), plot.best.Q = F, 
    methods = c('linear', 'lag', 'sat', 'lagsat')))) %>%
  rowwise %>%
  mutate(best.mod = mods$best.model, 
         mod.coef = list(coef(mods$best.model.contents[[1]])),
         B1 = ifelse('B1' %in% names(mod.coef), mod.coef['B1'], NA),
         B2 = ifelse('B2' %in% names(mod.coef), mod.coef['B2'], NA),
         a = ifelse(best.mod == 'gr', mod.coef['(Intercept)'], mod.coef['a']),
         b = ifelse(best.mod == 'gr', mod.coef['x'], mod.coef['b']),
         r.estim = b / a)

lag.sat.pars

mean.time.diff <- mono.data %>% group_by(strain, treat) %>% summarise(d = mean(diff(date.time))) %>% ungroup %>% summarise(d = mean(d)) %>% unlist %>% c

mono.data <- left_join(mono.data, lag.sat.pars) %>%
  mutate(cutoff.low =  B1 - 1.5 * mean.time.diff,#mean.time.diff/(5*b),
         cutoff.high = B2 + 3 * mean.time.diff,#mean.time.diff/(5*b),
         cutoff.low = ifelse(is.na(B1), 0, cutoff.low),
         cutoff.high = ifelse(is.na(B2), Inf, cutoff.high)) %>%
  ungroup %>%
  dplyr::filter(date.time > cutoff.low,
                date.time < cutoff.high)

```

```{r,plotting-raw-data}
# View raw data
ggplot(mono.data, aes(x = as.POSIXct(date), y = population.mean, col = strain)) + 
  facet_wrap(treat~., scales = 'free_y', nrow = 2, 
             labeller = label_wrap_gen(multi_line = FALSE)) +
    scale_y_log10(breaks = trans_breaks('log10', function(x){10^x}),
                labels = trans_format('log10', math_format(10^.x))) +
  geom_line(position = position_dodge(.5)) +
  geom_pointrange(aes(
    ymin = population.mean - population.sd, 
    ymax = population.mean + population.sd), 
    pch = 21, fill = 'white', position = position_dodge(.5)) +
  theme_bw() +
  labs(col = 'Strain', x = 'Time (Days)', y = 'Population Density') +
  theme(legend.position = 'bottom')
```


## Lotka-Volterra model fitting using GauseR
We now fit a Lotka-Volterra population model to the data, to determine the intrinsic growth rates and self-interaction parameters ($r$ and $a$, respectively).

```{r, fitting-parameters, fig.width=10, fig.height=8, fig.cap="Prediction of fitted Lotka-Volterra model"}
# model fitting
mod <- mono.data %>% 
  group_by(strain, treat) %>%
  summarise(
    gause.fit = list(gause_wrapper( # gauseR function
      time = date.time, # sampling intervals
      species = population.mean, # population densities
      keeptimes = TRUE, # return prediction at the same time-steps
      r_starting = ifelse(first(r.estim) < 0, 1e-8, first(r.estim)), # initial guess for r
      A_starting = -1e-8, # initial guess for a
      doplot = FALSE))# don't plot the model fitting process
    ) #%>%
  # rowwise %>%
  # mutate(r = gause.fit$parameter_intervals[2,2])

# extract model prediction
pred.data <- mod %>%
  ungroup %>% # we no longer need to do group-wise functions
  rowwise %>%
  mutate(pred = list(as.data.frame(gause.fit$out))) %>% # extract the model predictions
  dplyr::select(strain, treat, pred) %>% # remove unneeded columns
  unnest(pred) %>% # split prediction from nested lists into real columns
  rename(date.time = time, prediction = `1`) # rename columns

out.data <- pred.data %>% 
  full_join(mono.data) %>%
  mutate(d.pred = abs(prediction - population.mean))

# plot
ggplot(out.data, aes(x = date.time, col = strain)) + 
  facet_wrap(strain~treat, scales = 'free_y', nrow = 4, 
             labeller = label_wrap_gen(multi_line = FALSE)) +
  #scale_y_log10() +
  geom_line(aes(y = prediction), lty = 2) +
  geom_pointrange(aes(y = population.mean, ymin = population.mean - population.sd, 
                      ymax = population.mean + population.sd), 
                  pch = 21, fill = 'white') +
  theme_bw() +
  labs(col = 'Strain', x = 'Time (hours)', y = 'Population Density') +
  theme(legend.position = 'bottom')
```

### Accuracy
We'd like to see how accurate the model-fitting procedure is using the coefficient of determination... ($R^2$)

```{r, coefficient-of-determination, fig.width=7, fig.height=3, fig.cap="Accuracy of model-fitting procedure"}
# Data frame with CoD
CoD.data <- out.data %>%
  mutate(difference = population.mean - prediction,
         diff.squared = difference^2) %>%
  group_by(strain, treat) %>%
  summarise(RSS = sum(diff.squared),
            TSS = sum((population.mean - mean(population.mean))^2),
            R2 = 1 - (RSS / TSS),
            pop.range = population.mean %>% range %>% diff,
            population = mean(population.mean))

# plot
ggplot(CoD.data, aes(x = strain, y = R2, pch = treat, col = strain)) +
  facet_grid(.~treat) + 
  geom_point(size = 2, fill = 'white') +
  theme_bw() +
  labs(x = 'Coefficient of variation (CV) of trait values', 
       y = 'Coefficient of determination\nof model fit',
       pch = 'Treatment', col = 'Strain') +
  scale_shape_manual(values = 21:24) +
  theme(strip.placement = "outside", 
        strip.background = element_rect(fill = "white", linetype = 0), 
        axis.title.x = element_blank(), 
        legend.position = 'bottom')
```

We can see that the model fit is mostly good except for strain 2434 in atrazine and atrazine + temperature conditions.

### Accuracy and traits
What if the loss of model accuracy is due to functional trait changes? Let's first view the relevant ones...

```{r, d-traits-vs-acc}

# sd.traits <- out.data %>% 
#   group_by(strain, treat) %>%
#   summarise(across(FSC.HLin.mean:NIR.R.HLin.mean, sd)) %>% 
#   left_join(CoD.data) %>% 
#   pivot_longer(FSC.HLin.mean:NIR.R.HLin.mean, names_to = 'channel', values_to = 'SD') %>%
#   rowwise %>%
#   mutate(channel = strsplit(channel, '.HLin')[[1]][1])

# try with Euclidean distance?
sd.traits <- out.data %>% 
  group_by(strain, treat) %>%
  dplyr::select(date.time, contains('HLin.mean')) %>%
  dplyr::filter(row_number() %in% c(1, n())) %>% 
  summarise(across(FSC.HLin.mean:NIR.R.HLin.mean, function(i) {
    abs(i[2] - i[1]) / sum(i)
    #dist(i)[1] / sum(i)
    })) %>%
  pivot_longer(FSC.HLin.mean:NIR.R.HLin.mean) %>%
  group_by(strain, treat) %>%
  summarise(mean.d.trait = mean(value)) %>% 
  left_join(CoD.data)

ggplot(sd.traits, 
       aes(x = mean.d.trait, y = R2, col = treat)) +
  geom_point(aes(pch = strain)) +
  geom_smooth(method = 'lm', se = FALSE) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw() + 
  labs(y = expression('R'^2),
       x = 'Mean proportional change between initial & final traits',
       col = 'Treatment',
       pch = 'Strain',
       title = 'Mean proportional trait change vs model fitting accuracy')

# we have an issue here, since trait values change even while in lag phase...

```

How do they change over time?

```{r,traits-over-time, warning=FALSE, fig.width=8, fig.height=9, fig.cap="Changes in strain traits over time"}
# trait changes
trait.time <- out.data %>%
  dplyr::select(-contains('GRN.B'), -contains('NIR.B'), -contains('NIR.R')) %>%
  pivot_longer(contains('HLin'), names_to = c('channel', 'metric'), names_pattern = '(.*).HLin.(.*)') %>%
  pivot_wider(names_from = metric, values_from = value)
  
# plot
ggplot(trait.time, aes(x = date.time, y = mean, col = strain, pch = treat)) + 
  facet_grid(channel~strain, scales = 'free_y') +
  geom_line() +
  geom_pointrange(aes(ymax = mean + sd, ymin = mean - sd), fill = 'white') +
  theme_bw() +
  scale_shape_manual(values = c(21, 22, 23, 24)) +
  labs(col = 'Strain', x = 'Time (hours)', y = 'Trait value') +
  theme(legend.position = 'bottom')
```

Do the trait values predict per-capita growth?

```{r, traits-percap}

traits.pcgr <- out.data %>%
  dplyr::select(-contains('.sd')) %>%
  group_by(strain, treat) %>%
  mutate(pcgr = percap_growth(population.mean, 
                              dplyr::lag(population.mean), 
                              date.time - dplyr::lag(date.time))) %>% 
  pivot_longer(FSC.HLin.mean:NIR.R.HLin.mean, names_to = 'Channel') %>%
  rowwise %>%
  mutate(Channel = strsplit(Channel, '.HLin')[[1]][1]) %>%
  drop_na(pcgr)

# actually we want to pick channels which are not that different between strains ?

ggplot(traits.pcgr, 
       aes(x = value, y = pcgr)) +
  facet_wrap(.~Channel, scales = 'free', labeller = 'label_both') +
  geom_smooth(method = 'glm', formula = y ~ poly(x, 2), se = T) +
  geom_hline(yintercept = 0) +
  geom_point(aes(col = treat, pch = strain)) +
  theme_bw() +
  labs(y = 'Per-capita growth rate',
       x = 'Trait value')

```

How much does each trait change?

```{r, d-traits, fig.width=8, fig.height=6, fig.cap="Amount of trait change between beginning and end of experiment"}
# trait CV
trait.change <- trait.time %>%
  group_by(strain, treat, channel) %>%
  summarise(change = (first(mean) - last(mean)) / first(mean))#sd(value) / mean(value) * 100)

# plot
ggplot(trait.change, aes(x = strain, y = change, col = strain, pch = treat)) +
  facet_wrap(.~channel) +
  geom_hline(lty = 2, yintercept = 0) +
  geom_point(size = 2, position = position_dodge(width = .5)) +
  theme_bw() +
  scale_shape_manual(values = 21:24) +
  theme(legend.position = 'bottom')
```

How is this trait CV correlated to the coefficient of determination

```{r, d-traits-R2, fig.width=8, fig.height=6, fig.cap="Model fitting accuracy as a function of the amount of trait change"}
# merge
CoD.traits <- trait.change %>%
  left_join(CoD.data)

# plot
ggplot(CoD.traits, aes(x = change, y = R2, col = strain, pch = treat)) +
  facet_wrap(.~channel) +
  geom_vline(xintercept = 0, lty = 2) +
  geom_point(size = 2, fill = 'white') +
  theme_bw() +
  scale_shape_manual(values = 21:24) +
  #scale_y_continuous(limits = c(.75, 1)) +
  theme(legend.position = 'bottom')
```

This is pants.

# Do effects on traits predict growth
We have the strain growth parameters.

```{r, growth-par, fig.width=8, fig.height=3, fig.cap="Fitted model parameters"}
# extract parameters
mod.df <- mod %>% 
  rowwise %>%
  mutate(r = gause.fit$parameter_intervals['r','mu'],
         aii = gause.fit$parameter_intervals['a','mu']) %>%
  pivot_longer(r:aii, names_to = 'parameter')

# plot
ggplot(mod.df, aes(x = strain, y = value, col = strain, pch = treat)) +
  facet_wrap(parameter~., scales = 'free_y') +
  geom_point(size = 2, fill = 'white', 
             position = position_dodge(.5)) +
  scale_shape_manual(values = 21:24) +
  theme_bw() +
  theme(legend.position = 'bottom')
```


```{r, traits-growth-par, fig.cap="Fitted model parameters as a function of the mean trait value"}
# data framing
mod.traits <- trait.time %>% 
  group_by(strain, treat, channel) %>%
  summarise(channel.value = mean(mean),
            channel.sd = mean(sd)) %>%
  left_join(mod.df) %>%
  pivot_wider(names_from = parameter, values_from = value)

# intrinsic growth rate
ggplot(mod.traits, aes(x = channel.value, y = r)) +
  facet_wrap(~channel, scales = 'free') +
  scale_y_log10(breaks = trans_breaks('log10', function(x){10^x}),
                labels = trans_format('log10', scales::math_format(10^.x))) + 
  geom_smooth(method = 'lm') +
  geom_point(aes(col = treat, pch = strain),
             size = 2, fill = 'white') +
  scale_shape_manual(values = 21:24) +
  theme_bw() +
  labs(x = 'Trait mean value', 
       y = expression(paste('Intrinsic growth rate (', mu['i'], ')'))) +
  theme(legend.position = 'bottom')

# density-dependence
ggplot(mod.traits, aes(x = channel.sd, y = -aii)) +
  facet_wrap(~channel, scales = 'free') +
  scale_y_log10(breaks = trans_breaks('log10', function(x){10^x}),
                labels = trans_format('log10', scales::math_format(10^.x))) + 
  geom_smooth(method = 'glm') +
  geom_point(aes(col = treat, pch = strain),
             size = 2, fill = 'white') +
  scale_shape_manual(values = 21:24) +
  theme_bw() +
  labs(x = 'Trait standard deviation (SD)', 
       y = expression(paste('Self-limitation (', alpha['11'], ')'))) +
  theme(legend.position = 'bottom')


```

# Do certain traits predict sensitivity to atrazine exposure?

We want the starting trait value on the x vs the change in a, K, r on the y axis

<!-- ```{r, diff-traits-growth-par} -->

<!-- # FIXME -->

<!-- # calculate change in parameters -->
<!-- traits.start <- trait.time %>% -->
<!--   dplyr::filter(treat == 'Control', date == min(date)) %>% -->
<!--   rename(Channel.value = value) %>% -->
<!--   dplyr::select(-treat, -date.time, -prediction, -date, -population, -pop.vec, -d.pred) -->

<!-- dpar <- mod.df %>% -->
<!--   group_by(strain, Parameter) %>% -->
<!--   summarise(Parameter.control = value[treat == 'C'], -->
<!--             Parameter.atr = value[treat == 'A'], -->
<!--             Parameter.change = (Parameter.atr)/Parameter.control) %>% -->
<!--   full_join(traits.start) -->

<!-- # plot  -->
<!-- ggplot(dpar, aes(x = Channel.value, y = Parameter.change, col = strain)) + -->
<!--   facet_grid(Parameter~Channel, scale = 'free') + -->
<!--   geom_hline(yintercept = 1, lty = 2) + -->
<!--   geom_point(size = 2, fill = 'white', pch = 21) + -->
<!--   theme_bw() -->
<!-- ``` -->


<!-- # Does the presence of competitors affect trait changes? -->
<!-- ## Strain pairs -->
<!-- ```{r, pairwise-plotting-raw-data, fig.width=5, fig.height=10} -->
<!-- # import data -->
<!-- pair.df <- readRDS('C:/Users/mholmes/Documents UNamur/Lab work/03_Experiments/12_2022-01-14_DIVERCE/pair_data_clean.RData') %>% -->
<!--   group_by(strains, treat) %>% -->
<!--   arrange(date.time) %>% -->
<!--   mutate(date.time = as.numeric(date.time - min(date.time)) / 3600) %>% #in hours -->
<!--   dplyr::filter(treat == 'T' | treat == 'AT') %>% # remove extra treatments -->
<!--   mutate(treat = recode(treat, 'T' = 'Control', 'AT' = 'Atrazine'), # Rename treatment levels -->
<!--          treat = as.factor(treat), -->
<!--          treat = relevel(treat, 'Control', 'Atrazine')) -->

<!-- # plot -->
<!-- ggplot(pair.df, aes(x = date.time)) +  -->
<!--   facet_wrap(strains~treat, scales = 'free', ncol = 2,  -->
<!--              labeller = label_wrap_gen(multi_line=FALSE)) + -->
<!--   geom_line(aes(y = population.1, col = strain.1)) + -->
<!--   geom_line(aes(y = population.2, col = strain.2)) + -->
<!--   geom_point(aes(y = population.1, col = strain.1), -->
<!--              pch = 21, fill = 'white', size = 2) + -->
<!--   geom_point(aes(y = population.2, col = strain.2), -->
<!--              pch = 21, fill = 'white', size = 2) + -->
<!--   theme_bw() + -->
<!--   theme(legend.position = 'bottom') -->
<!-- ``` -->


<!-- ## Lotka-Volterra model fitting using GauseR -->
<!-- We now fit a Lotka-Volterra population model to the data, to determine the intrinsic growth rates and self/inter-interaction parameters ($r$ and $a$, respectively). -->

<!-- ```{r, pair-fitting-parameters, fig.width=5, fig.height=10, fig.cap="Prediction of fitted Lotka-Volterra model"} -->

<!-- ## fit -->
<!-- pair.mod <- pair.df %>% -->
<!--   ungroup %>% -->
<!--   arrange(strains, treat, date.time) %>% -->
<!--   group_by(strains, treat) %>% -->
<!--   summarise(gause.fit = tryCatch(list(gause_wrapper( -->
<!--         r_starting = rep(5e-2, 2), -->
<!--         A_starting = rep(-5e-7, 4), -->
<!--         time = date.time, -->
<!--         species = data.frame(population.1, population.2), -->
<!--         doplot = FALSE, -->
<!--         keeptimes = TRUE)), -->
<!--       error = function(e) {NULL}), -->
<!--     strain.1 = strain.1[1], -->
<!--     strain.2 = strain.2[1]) -->

<!-- pair.pred <- pair.mod %>%  -->
<!--   ungroup %>%  -->
<!--   rowwise %>%  -->
<!--   dplyr::filter(length(gause.fit) != 1) %>% -->
<!--   mutate(pred = list(as.data.frame(gause.fit$out))) %>% -->
<!--   dplyr::select(strains, treat, pred) %>%  -->
<!--   unnest(pred) %>% -->
<!--   rename(date.time = time, prediction.1 = population.1, prediction.2 = population.2) -->

<!-- pair.out <- pair.pred %>%  -->
<!--   full_join(pair.df) %>% -->
<!--   mutate(d.pred.1 = abs(prediction.1 - population.1), -->
<!--          d.pred.2 = abs(prediction.2 - population.2), -->
<!--          treat = recode(treat, 'T' = 'Control', 'AT' = 'Atrazine'), -->
<!--          treat = as.factor(treat), -->
<!--          treat = relevel(treat, 'Control', 'Atrazine')) -->

<!-- # plot -->
<!-- ggplot(pair.out, aes(x = date.time, pch = treat)) +  -->
<!--   facet_wrap(strains~treat, scales = 'free', ncol = 2, -->
<!--              labeller = label_wrap_gen(multi_line=FALSE)) + -->
<!--   geom_line(aes(y = prediction.1, col = strain.1), lty = 2) + -->
<!--   geom_line(aes(y = prediction.2, col = strain.2), lty = 2) + -->
<!--   geom_point(aes(y = population.1, col = strain.1), -->
<!--              fill = 'white', size = 2) + -->
<!--   geom_point(aes(y = population.2, col = strain.2),  -->
<!--              fill = 'white', size = 2) + -->
<!--   theme_bw() + -->
<!--   scale_shape_manual(values = c(21, 24))+ -->
<!--   labs(col = 'Strain', x = 'Time (hours)', y = 'Population Density') + -->
<!--   theme(legend.position = 'bottom') -->

<!-- ``` -->


<!-- ### Accuracy -->
<!-- We'd like to see how accurate the model-fitting procedure is using the coefficient of determination... ($R^2$) -->

<!-- ```{r, pair-coefficient-of-determination, fig.width=5, fig.height=4, fig.cap="Accuracy of model-fitting procedure"} -->
<!-- # Data frame with CoD -->
<!-- pair.CoD <- pair.out %>% -->
<!--   mutate(difference = (population.1 - prediction.1) + (population.2 - prediction.2), -->
<!--          diff.squared = difference^2) %>% -->
<!--   group_by(strains, treat) %>% -->
<!--   summarise(RSS = sum(diff.squared), -->
<!--             TSS = sum((population.1 - mean(population.1))^2,  -->
<!--                       (population.2 - mean(population.2))^2), -->
<!--             R2 = 1 - (RSS / TSS), -->
<!--             pop.1.range = population.1 %>% range %>% diff, -->
<!--             pop.2.range = population.2 %>% range %>% diff, -->
<!--             population.1 = mean(population.1), -->
<!--             population.2 = mean(population.2)) -->

<!-- # plot -->
<!-- ggplot(pair.CoD, aes(x = strains, y = R2, pch = treat, col = strains)) + -->
<!--   facet_grid(.~treat) +  -->
<!--   geom_point(size = 2, fill = 'white') + -->
<!--   theme_bw() + -->
<!--   scale_shape_manual(values = c(21, 24)) + -->
<!--   labs(x = 'Coefficient of variation (CV) of trait values',  -->
<!--        y = 'Coefficient of determination\nof model fit', -->
<!--        pch = 'Treatment', col = 'Strain') + -->
<!--   theme(strip.placement = "outside",  -->
<!--         axis.text.x = element_text(angle = 90, vjust = 0.5), -->
<!--         strip.background = element_rect(fill = "white", linetype = 0),  -->
<!--         axis.title.x = element_blank(),  -->
<!--         legend.position = 'bottom', legend.box = 'vertical') -->
<!-- ``` -->


<!-- ### Accuracy and traits -->
<!-- What if the loss of model accuracy is due to functional trait changes? Let's first view the relevant ones... How do they change over time? -->

<!-- ```{r} -->
<!-- test.pair <- pair.out %>%  -->
<!--   dplyr::filter(strain.1 == '2375') %>% -->
<!--   mutate(FSC.HLin.1 = FSC.HLin$mean.1, -->
<!--          FSC.HLin.2 = FSC.HLin$mean.2)  -->

<!-- ggplot(test.pair, aes(x = date.time, pch = treat)) + -->
<!--   facet_grid(strain.2~treat) + -->
<!--   geom_point(aes(y = FSC.HLin.1, col = 'strain_1')) + -->
<!--   geom_point(aes(y = FSC.HLin.2, col = 'strain_2')) + -->
<!--   theme_bw() -->
<!-- ``` -->


<!-- ```{r, pair-traits-over-time, warning=FALSE, fig.width=8, fig.height=10, fig.cap="Changes in strain traits over time"} -->
<!-- # trait changes -->
<!-- pair.trait.time <- pair.out %>% -->
<!--   dplyr::select(-GRN.B.HLin, -NIR.B.HLin, -NIR.R.HLin) %>% -->
<!--   mutate(trait.1 = across(FSC.HLin:RED.R.HLin, .fns = function(i) {i$mean.1}), -->
<!--          trait.2 = across(FSC.HLin:RED.R.HLin, .fns = function(i) {i$mean.2})) %>% -->
<!--   dplyr::select(-(FSC.HLin:RED.R.HLin)) %>% -->
<!--   unnest_wider(trait.1:trait.2, names_sep = '_') -->

<!-- # strain 2375 vs competitor data frame -->
<!-- pair.2375 <- pair.trait.time %>% -->
<!--   dplyr::filter(strain.1 == '2375' | strain.2 == '2375') %>% -->
<!--   mutate(strain = 2375, -->
<!--          competitor = ifelse(strain.1 == '2375', strain.2, strain.1), -->
<!--          d.pred = ifelse(strain.1 == '2375', d.pred.1, d.pred.2), -->
<!--          FSC.HLin = ifelse(strain.1 == '2375', trait.1_FSC.HLin, trait.2_FSC.HLin), -->
<!--          SSC.HLin = ifelse(strain.1 == '2375', trait.1_SSC.HLin, trait.2_SSC.HLin), -->
<!--          YEL.B.HLin = ifelse(strain.1 == '2375', trait.1_YEL.B.HLin, trait.2_YEL.B.HLin), -->
<!--          RED.B.HLin = ifelse(strain.1 == '2375', trait.1_RED.B.HLin, trait.2_RED.B.HLin), -->
<!--          RED.R.HLin = ifelse(strain.1 == '2375', trait.1_RED.R.HLin, trait.2_RED.R.HLin)) %>% -->
<!--   dplyr::select(strains, treat, date.time, strain:RED.R.HLin) -->

<!-- # strain 2383 vs competitor data frame -->
<!-- pair.2383 <- pair.trait.time %>% -->
<!--   dplyr::filter(strain.1 == '2383' | strain.2 == '2383') %>% -->
<!--   mutate(strain = 2383, -->
<!--          competitor = ifelse(strain.1 == '2383', strain.2, strain.1), -->
<!--          d.pred = ifelse(strain.1 == '2383', d.pred.1, d.pred.2), -->
<!--          FSC.HLin = ifelse(strain.1 == '2383', trait.1_FSC.HLin, trait.2_FSC.HLin), -->
<!--          SSC.HLin = ifelse(strain.1 == '2383', trait.1_SSC.HLin, trait.2_SSC.HLin), -->
<!--          YEL.B.HLin = ifelse(strain.1 == '2383', trait.1_YEL.B.HLin, trait.2_YEL.B.HLin), -->
<!--          RED.B.HLin = ifelse(strain.1 == '2383', trait.1_RED.B.HLin, trait.2_RED.B.HLin), -->
<!--          RED.R.HLin = ifelse(strain.1 == '2383', trait.1_RED.R.HLin, trait.2_RED.R.HLin)) %>% -->
<!--   dplyr::select(strains, treat, date.time, strain:RED.R.HLin) -->

<!-- # strain 2434 vs competitor data frame -->
<!-- pair.2434 <- pair.trait.time %>% -->
<!--   dplyr::filter(strain.1 == '2434' | strain.2 == '2434') %>% -->
<!--   mutate(strain = 2434, -->
<!--          competitor = ifelse(strain.1 == '2434', strain.2, strain.1), -->
<!--          d.pred = ifelse(strain.1 == '2434', d.pred.1, d.pred.2), -->
<!--          FSC.HLin = ifelse(strain.1 == '2434', trait.1_FSC.HLin, trait.2_FSC.HLin), -->
<!--          SSC.HLin = ifelse(strain.1 == '2434', trait.1_SSC.HLin, trait.2_SSC.HLin), -->
<!--          YEL.B.HLin = ifelse(strain.1 == '2434', trait.1_YEL.B.HLin, trait.2_YEL.B.HLin), -->
<!--          RED.B.HLin = ifelse(strain.1 == '2434', trait.1_RED.B.HLin, trait.2_RED.B.HLin), -->
<!--          RED.R.HLin = ifelse(strain.1 == '2434', trait.1_RED.R.HLin, trait.2_RED.R.HLin)) %>% -->
<!--   dplyr::select(strains, treat, date.time, strain:RED.R.HLin) -->

<!-- # strain 2524 vs competitor data frame -->
<!-- pair.2524 <- pair.trait.time %>% -->
<!--   dplyr::filter(strain.1 == '2524' | strain.2 == '2524') %>% -->
<!--   mutate(strain = 2524, -->
<!--          competitor = ifelse(strain.1 == '2524', strain.2, strain.1), -->
<!--          d.pred = ifelse(strain.1 == '2524', d.pred.1, d.pred.2), -->
<!--          FSC.HLin = ifelse(strain.1 == '2524', trait.1_FSC.HLin, trait.2_FSC.HLin), -->
<!--          SSC.HLin = ifelse(strain.1 == '2524', trait.1_SSC.HLin, trait.2_SSC.HLin), -->
<!--          YEL.B.HLin = ifelse(strain.1 == '2524', trait.1_YEL.B.HLin, trait.2_YEL.B.HLin), -->
<!--          RED.B.HLin = ifelse(strain.1 == '2524', trait.1_RED.B.HLin, trait.2_RED.B.HLin), -->
<!--          RED.R.HLin = ifelse(strain.1 == '2524', trait.1_RED.R.HLin, trait.2_RED.R.HLin)) %>% -->
<!--   dplyr::select(strains, treat, date.time, strain:RED.R.HLin) -->

<!-- # rebind -->
<!-- pair.trait.time.2 <- rbind(pair.2375, pair.2383, pair.2434, pair.2524) %>% -->
<!--   pivot_longer(FSC.HLin:RED.R.HLin, names_to = 'Channel') -->

<!-- # plot -->
<!-- ggplot(pair.trait.time.2, aes(x = date.time, y = value, col = competitor, -->
<!--            pch = treat)) + -->
<!--   facet_grid(Channel~strain, scales = 'free_y') + -->
<!--   geom_line() + -->
<!--   geom_point(size = 2) + -->
<!--   theme_bw() + -->
<!--   labs(col = 'Competing strain', x = 'Time (hours)', y = 'Trait value', pch = 'Treatment') + -->
<!--   theme(legend.position = 'bottom') -->
<!-- ``` -->

<!-- How much does each trait change? -->

<!-- ```{r, pair-d-traits, fig.width=8, fig.height=6, fig.cap="Amount of trait change between beginning and end of experiment"} -->
<!-- # trait CV -->
<!-- pair.trait.change <- pair.trait.time.2 %>% -->
<!--   group_by(strain, competitor, treat, Channel) %>% -->
<!--   summarise(strains = unique(strains),  -->
<!--             change = (first(value) - last(value)) / first(value))#sd(value) / mean(value) * 100) -->

<!-- # plot -->
<!-- ggplot(pair.trait.change,  -->
<!--        aes(x = as.factor(strain), y = change, col = competitor, pch = treat)) + -->
<!--   facet_wrap(.~Channel) + -->
<!--   geom_hline(lty = 2, yintercept = 0)+ -->
<!--   geom_point(size = 2, fill = 'white', position = position_dodge(width = 0.25)) + -->
<!--   theme_bw() + -->
<!--   scale_shape_manual(values = c(21, 24)) + -->
<!--   theme(legend.position = 'bottom') -->
<!-- ``` -->

<!-- # **COMPLETED UP UNTIL HERE** -->

<!-- How is this trait CV correlated to the coefficient of determination -->

<!-- ```{r, pair-d-traits-R2, fig.width=8, fig.height=6, fig.cap="Model fitting accuracy as a function of the amount of trait change"} -->
<!-- # merge -->
<!-- pair.CoD.traits <- pair.trait.change %>% -->
<!--   left_join(pair.CoD) -->

<!-- # plot -->
<!-- ggplot(pair.CoD.traits, aes(x = change, y = R2, col = competitor, pch = treat)) + -->
<!--   facet_grid(strain~Channel) + -->
<!--   geom_vline(xintercept = 0, lty = 2) + -->
<!--   geom_point(size = 2, fill = 'white') + -->
<!--   theme_bw() + -->
<!--   scale_shape_manual(values = c(21, 24)) + -->
<!--   scale_y_continuous(limits = c(.75, 1)) + -->
<!--   theme(legend.position = 'bottom') -->
<!-- ``` -->

<!-- This is pants. -->

<!-- # **COMPLETED UP UNTIL HERE** -->

<!-- # Do effects on traits predict growth -->
<!-- We have the strain growth parameters. -->

<!-- ```{r, pair-growth-par, fig.width=8, fig.height=8, fig.cap="Fitted model parameters"} -->
<!-- # extract parameters -->
<!-- pair.mod.df <- pair.mod %>%  -->
<!--   rowwise %>% -->
<!--   mutate(r1 = gause.fit$parameter_intervals['r1','mu'], -->
<!--          r2 = gause.fit$parameter_intervals['r2','mu'], -->
<!--          a11 = gause.fit$parameter_intervals['a11','mu'], -->
<!--          a12 = gause.fit$parameter_intervals['a12','mu'], -->
<!--          a21 = gause.fit$parameter_intervals['a21','mu'], -->
<!--          a22 = gause.fit$parameter_intervals['a22','mu'], -->
<!--          #K = -r / aii, -->
<!--          ) %>% -->
<!--   #pivot_longer(r:K, names_to = 'Parameter') -->
<!--   pivot_longer(r1:a22, names_to = 'Parameter') -->

<!-- # plot -->
<!-- ggplot(pair.mod.df, aes(x = strain.1, y = value, col = strain.2)) + -->
<!--   facet_grid(Parameter~strain.2, scales = 'free_y') + -->
<!--   #scale_y_continuous(trans = 'pseudo_log') + -->
<!--   geom_path(lty = 2) + -->
<!--   geom_point(aes(pch = treat), size = 2) + -->
<!--   theme_bw() + -->
<!--   theme(legend.position = 'bottom') -->
<!-- ``` -->


<!-- ```{r, pair-traits-growth-par, fig.width=8, fig.height=8, fig.cap="Fitted model parameters as a function of the mean trait value"} -->
<!-- pair.mod.traits <- pair.trait.time.2 %>%  -->
<!--   group_by(strains, treat, Channel) %>% -->
<!--   rename(Channel.value = value) %>% -->
<!--   summarise(Channel.value = mean(Channel.value)) %>% -->
<!--   left_join(pair.mod.df) %>% -->
<!--   rename(Parameter.value = value) -->

<!-- # plot -->
<!-- ggplot(pair.mod.traits, aes(x = Channel.value, y = Parameter.value, col = strains)) + -->
<!--   facet_grid(Parameter~Channel, scales = 'free') + -->
<!--   scale_y_continuous(trans = 'pseudo_log') + -->
<!--   geom_path(lty = 2) + -->
<!--   geom_point(aes(pch = treat), size = 2) + -->
<!--   theme_bw() + -->
<!--   theme(legend.position = 'bottom') -->
<!-- ``` -->

<!-- Hey look at 2375, its density dependence never changes due to stress, maybe that's a nice one to keep in mind for later comparison. -->

<!-- # Do certain traits predict sensitivity to atrazine exposure? -->

<!-- We want the starting trait value on the x vs the change in a, K, r on the y axis -->

<!-- ```{r, pair-diff-traits-growth-par} -->
<!-- # calculate change in parameters -->
<!-- traits.start <- trait.time %>% -->
<!--   dplyr::filter(treat == 'Control', date == min(date)) %>% -->
<!--   rename(Channel.value = value) %>% -->
<!--   dplyr::select(-treat, -date.time, -prediction, -date, -population, -pop.vec, -d.pred) -->

<!-- dpar <- mod.df %>% -->
<!--   group_by(strain, Parameter) %>% -->
<!--   summarise(Parameter.control = value[treat == 'Control'], -->
<!--             Parameter.atr  = value[treat == 'Atrazine'], -->
<!--             Parameter.change = (Parameter.atr)/Parameter.control) %>% -->
<!--   full_join(traits.start) -->

<!-- # plot  -->
<!-- ggplot(dpar, aes(x = Channel.value, y = Parameter.change, col = strain)) + -->
<!--   facet_grid(Parameter~Channel, scale = 'free') + -->
<!--   #scale_y_continuous(trans = 'pseudo_log') + -->
<!--   geom_hline(yintercept = 1, lty = 2) + -->
<!--   geom_point() + -->
<!--   theme_bw() -->
<!-- ``` -->

Let's try without the outlier

